---
title: "Tutorial for farsutil package"
author: "Lukasz Krainski"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial for farsutil package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(farsutil)
library(dplyr)
```

## Exported functions

- `fars_read()`
- `make_filename()`
- `fars_read_years()`
- `fars_summarize_years()`
- `fars_map_state()`

## Reading FARS data files

The function `fars_read()` read data from provided filename (returns a `tibble`):

```{r fars_read_example}
filename <- system.file("data/accident_2013.csv.bz2", package = "farsutil")
fars_read(filename)
```

It utilizes `make_filename()` auxilary function.

```{r fars_makefile}
make_filename(2015)
```

To read multiple years at once use `fars_read_years()` function.

```{r fars_read_years}
setwd(system.file("data", package = "farsutil"))
fars_read_years(2013:2015)
```

## Summarize number of accidents

`fars_summarize_years()` returns a pivot table with number of accidents per year and month for provided years. Data must be located in the working directory.

```{r fars_summarize_years}
setwd(system.file("data", package = "farsutil"))
fars_summarize_years(2013:2015)
```

## Plot accident locations

`fars_map_state()` returns a plot with accident locations for the given state.

```{r fars_map_state}
setwd(system.file("data", package = "farsutil"))
fars_map_state(45, 2015)
```

